\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-20 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript,
``Looking forwards and backwards: dynamics and
genealogies of locally regulated populations''.
Thanks to the reviewer's comments
(as well as separate comments from Mattias Birkner),
we think the manuscript has been improved and clarified in a number of places,
a few (happily, relatively minor) errors and omissions
have been fixed,
and the manuscript is generally improved.
We hope you agree.

We have submitted a clean version of the revised paper and
the response to the reviewer as separate PDFs;
line numbers in this response refer to the ``clean'' version.
Also, we have submitted a single PDF
with changes to the previous version shown (with latexdiff)
and a version of the response to reviewer at the end of the document.
In this combined ``diff''document, line numbers (and hyperlinks)
in the reviewer responses are consistent within the document.

@@

\begin{point}{}
    Figure \ref{fig:super_vs_det_2d}: What is $N$, the time and other parameters (e.g. $\rho_r$ , $\rho_\gamma$) in the simulations? Same
for other figures. Please at least specify the range of values.
\end{point}

\reply{
    We've added some more detail to the caption.
}

\begin{point}{\revref}
% P12 paragraph after Corollary 2.11:
A more precise reference, or explanation for why
uniqueness holds for this special case would be helpful, as the references you ave are all
quite extensive.
\end{point}

\reply{
    Good suggestion -- this was rather cryptic.
    We've expanded on what we meant here \revref.
}

@@

\reply{
    Good point! Done.
}

\begin{point}{}
Remark \ref{remark_on_nonlocal_to_local}: Can you explain a bit more why it will be necessary to retain averaging of r?
\end{point}

\reply{
    Good idea; done \revref.
}

\begin{point}{}
% P16: Second paragraph:
The notation $L_t$ seems to first appear here and was not yet defined.
\end{point}

@@

\reply{
    Ah, true - now we introduce it (albeit still informally) earlier \llname{Linf_mention}.
}

\begin{point}{}
% P19-20:
    The equations for the lineage dynamics for the three examples [i.e., Fisher-KPP, Allen-Cahn, and Porous Medium] are interesting.
Which is/are new in this paper, and which was/were known before (and related references)?
\end{point}

\reply{
    As far as we know, the PME case is new,
    but there's been a fair amount of work on FKPP-like
    cases covering the other two cases.
    We've put a note about this, 
    and some pointers to this literature \revref.
}

@@

\begin{point}{\revref}
specify ``this process'' in the line below the first displayed equation in
Section 5.1.
\end{point}

\reply{
    Done \revref.
}

\begin{point}{\revref}
Paragraph starting with ``Levels never cross below 0'': Should it be instead ``Levels never
go above $N$''? If so, what particles are regarded as dead? Perhaps those exit the interval
$[0, N]$? Note that equation \eqref{eqn:dot_u} suggests that levels can now go above $N$.
\end{point}

\reply{
    This was correct as written;
    we've rephrased it to be more clear \revref.
}

@@

\begin{point}{\revref}
% P32 third line in Section 5.3:
    Pointwise in time?
\end{point}

\reply{
    We mean for each $\omega$; now clarified \revref.
}

\begin{point}{\revref}
Proposition \ref{prop:limiting_construction}: In ``For any solution to the equation above'', what is a ``solution''? Is it
a solution for all time? Also, be specific to ``the equations above'' by labeling and referring
to the precise equations including the unlabeled displayed equation below \eqref{eqn:limiting_construction}.
\end{point}

\reply{
    Good suggestion; we've clarified, see \revref{} and \llname{XU_clarification}.
}

@@

\begin{point}{}
The word ``Proof'' is missing in ``Proof of Proposition 5.6''. This word is also
missing in other places in Sections 6-8.
\end{point}

\reply{
    Whoops! That was due to a \LaTeX{} format change; now fixed.
}

\begin{point}{}
``Proof of Proposition \ref{prop:limiting_construction}'': Do you also need the unlabeled displayed equation below
\eqref{eqn:limiting_construction}? The solution is for all time.
\end{point}

\reply{
    We think this point was an unedited duplicate of point 1.24 above? (If not, we're not sure what else it's saying?)
}

@@

\begin{point}{\revref}
%    P48:
    The notation $\mathcal{O}(\epsilon)$ seems to first appear here. Please explain its meaning. Do they
depend on x?
\end{point}

\reply{
    We've made the bounds here (and below) more explicit.
    \revref
}

\begin{point}{\revref}
% P54: near the bottom:
    Can you be more specific what in paper \citet{barlow/jacka/yor:1986} is being used here?
\end{point}

\reply{
    It was Item (4.b'), Table 4.1, p. 162, again --
    we've modified this to hopefully 
    more clearly refer to the previous place we used the same argument
    \revref.
}

### 2023-12-20 forwards-and-backwards.tex
final plr edits

\comment{TODO: insert statement that one might be worried about $\eta_t$ being well-defined as a process, in e.g., Def'n 5.2 and following?}

@@

By replacing $\rho$ by
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, 
we can also scale the ``width'' of the 
region over which we measure local population density. 
When the population follows~(\ref{general deterministic limit}), 
we expect that if we take a second limit of $\epsilon \to 0$,
thus scaling the kernels appearing in $r$, $\gamma$, and $F$ and making interactions pointwise,
we should recover a nonlinear PDE.
We verify that this is indeed the case in two important examples:
a special case of the porous medium equation with a logistic growth term, in which
the limiting equation takes the form 
\begin{equation}
	\label{PME1}
	\partial_t \varphi = \Delta (\varphi^2)+\varphi(1-\varphi);
\end{equation}
and a wide class of semi-linear PDEs of the form 
\begin{equation}
\label{semilinear PDE}
\partial_t \varphi = \DG^*\varphi+ \varphi F(\varphi),
\end{equation}
which includes the Fisher-KPP equation and the Allen-Cahn equation.
Equations of this form have been studied extensively
in the context of spatial ecology
(see for instance \citet{lam2023introduction} and \citet{cantrell2004spatial})
and in many other fields;
for instance, \citet{ghosh2022emergent}
derive a stochastic version of~\eqref{semilinear PDE} to describe abundances of mutant bacteria strains
along the human gut,
while \citet{li/buenzli/simpson:2022} study the effects of nonlinear diffusion
on long-term survival of a lattice-based interacting particle system.
However, we do not study the effect of movement of adults,
which can additionally affect the limiting equations:
see for instance \citet{holmes/lewis/banks/veit:1994} or \citet{potts/borger:2023}.
% MORE DETAILS:
% \cite{li/buenzli/simpson:2022} write down an interacting particle
% system on a hexagonal lattice (with at most one individual per site) that incorporates an effect
% of local crowding on individual motion. Their equation
% corresponds to a reaction-diffusion equation with nonlinear diffusion and a 
% bistable reaction term of the form~(\ref{Birzu nonlinearity}) and they
% investigate the interplay between nonlinear diffusion and long-term 
% survival or extinction of the population.

@@

Since $F$ is bounded above,
the final assumption that $\mu_\theta \ge 0$ --
or, equivalently, that $F(x,\eta) \le \theta r(x,\eta) \gamma(x,\eta)$ --
will always be true, for instance, for large enough $\theta$ 
as long as $r$ and $\gamma$ are bounded away from zero.

@@

\begin{figure}
    \begin{center}
        % \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        % \includegraphics{figures/ex1b/fkpp_123.density.3.0}
        \includegraphics[width=2.5in]{figures/fig1a}
        \includegraphics[width=2.5in]{figures/fig1b}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha = \theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal ($q_\theta$) and interactions (here, only $\rho_F$)
        are Gaussian with standard deviation 1,
        and the equilibrium density (also, $N$) is 10 individuals per unit area.
        The remaining parameters are constant: $r \equiv \gamma \equiv 1$.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\citep{etheridge:2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness \revpoint{1}{12}
in this special case of the superprocess with nonlinear death rates \cite{dawson:1993}.
The Perkins stochastic calculus
(and its adaptation to a lookdown setting \cite{donnelly/kurtz:1999})
provides uniqueness for a different but related class of processes,
in which interactions affect the dispersal mechanism
(rather than reproduction) of the superprocess \cite{perkins:1992}.
We refer to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that these sorts of martingale problems are well-posed.

@@

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
$\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
Furthermore, suppose that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exists $\lambda\in (0,1)$
\begin{equation*}
\sup_\epsilon \int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx < \infty,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx < \infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}

@@

However, the \emph{lookdown construction} in Section~\ref{sec:lookdown}
does just this -- \llabel{whatis_lineage}
the construction enables us to recover information about ancestry
in the infinite population limit,
and thus give a concrete meaning to $(L_t)_{t \ge 0}$. \llabel{Linf_mention}
Roughly speaking,
each particle is assigned a unique ``level'' from $[0,\infty)$
that functions as a label and thus allows reconstruction of lineages.
The key to the approach is that levels are assigned in such a way as to be exchangeable, 
so that sampling a finite number, $k$ say, of
individuals from a given 
region is equivalent to looking at the individuals in that region with the $k$ lowest levels.
Moreover, as we pass to the infinite population limit,
the collection of (individual, level) pairs converges,
as we show in Theorem~\ref{thm:lookdown_convergence}.
See~\citet{etheridge/kurtz:2019} for an introduction to these ideas.
In particular, even in the infinite
population limit, we can sample an individual
from a region (it will be the individual in that region with the lowest level) 
and trace its line of descent. 
This will allow us to calculate, for each $x$ and $y\in\IR^d$, the proportion of
the population at location $x$ in the present day population that are descended from 
a parent who was at location $y$ at time $t$ in the past. To make
sense of this in our framework, in Section~\ref{sec:lineages_proof}, 
we justify a weak reformulation of this idea.

@@

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
with mean displacement in the direction of $\grad \log(\varphi_s \gamma) - \meanq$.
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

@@

%%%%%%%%%%

%%%%%%%%%%
\paragraph{Fisher-KPP equation:} \revpoint{1}{16}
Consider the classical Fisher-KPP equation,
\begin{align} \label{eqn:fkpp}
    \partial_t\varphi = \partial_{xx}\varphi + \varphi (1-\varphi) .
\end{align}
Even though we do not have an explicit formula for the wave shape in this case,
our methods provide information about ancestral lineages.
The equation has non-negative travelling wave solutions of 
speed $c$ for all $c \geq 2$, 
but, started from any compact perturbation of a Heaviside function, the 
solution will converge to the profile $w^F$ with the minimal wavespeed, $c=2$
\citep{kolmogorov/petrovsky/piscounov:1937,fife1977approach,bramson:1983}.
No matter what initial condition,
for any $t>0$ the support of the 
solution will be the whole real line. 
In this case, we must have $r = \gamma = 1$,
and $F(x, m) = 1 - m$ so $\mu_\theta(x, m) = 1 + (m-1)/\theta$.
By Corollary~\ref{cor:wavefront},
the generator of the motion of an ancestral lineage is
\begin{equation} \label{eqn:fkpp_generator}
    \Lgen_F  f
    =
    \partial_{xx}f + 2 \frac{\partial_xw^F}{w^F} \partial_x f + 2 \partial_xf .
\end{equation}
Near the tip of the wave (for $x$ large), $w^F(x) \sim e^{-x}$,
so \eqref{eqn:fkpp_generator} implies that
the motion of a lineage is close to unbiased Brownian motion.
On the other hand, in the ``bulk'', %for $x < 0$ 
a lineage behaves approximately as
Brownian motion with drift at rate two to the right.
This implies that
ancestral lineages are pushed into the tip of the wave,
and there is no stationary distribution,
so that long-term dynamics of genetic inheritance
depend on the part of the wave not well-approximated by a smooth profile,
in agreement with the previous results referred to in the Introduction.

@@

We now look at the second order term in~(\ref{generator prelimit}),
which will converge to the quadratic variation of the limiting process.
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\int
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
\eta(dx) .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $-f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
\begin{align*}
& G''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \int
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    \eta(dx)
    +
    \int
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    \eta(dx)
\right\} \\
&\qquad
= \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
	\int\left\{
		\gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta)
    \right\} \eta(dx) .
\end{align*}
Since $\int f^2(y) r(y, \eta) q_\theta(x, dy) \to f^2(x) r(x, \eta)$
and $r\gamma  + \mu_\theta = 2 r \gamma  - F / \theta \to 2 r \gamma$
as $\theta \to \infty$,
this converges to
\begin{align*}
\frac{\alpha}{2} G''(\langle f, \eta \rangle)
    \big\langle
	2 r(x, \eta) \gamma(x, \eta),
        \eta(dx)
    \big\rangle .
\end{align*}

@@

\paragraph{One-step convergence:} \label{sec:one_step_heuristics} \llabel{one_step_heuristics}
In order to pass directly to a classical PDE limit
in Theorem~\ref{thm:local_convergence}
we impose the stronger condition
that $\theta/N\epsilon^d\to 0$ and also require that
$\theta\epsilon^2\to\infty$. 
Recall that in this case, we take $\rho_F^\epsilon$ 
to be a symmetric Gaussian density with variance 
$\epsilon^2$. The condition $\theta\epsilon^2\to\infty$ 
ensures that $\epsilon^2$ is large enough relative to $1/\theta$
that the regularity gained by smoothing our population density by convolution with
$\rho_\epsilon$ is preserved under the dynamics dictated by $q_{\theta}$.
To understand the first condition, note that we are aiming to obtain a 
deterministic expression for the limiting population density. 
It is helpful to think
about a classical Wright-Fisher model (with no spatial structure and just two types, say). 
We know then that if the
timescale $\theta$ is on the same order as population size $N$, we see stochastic
fluctuations in the frequencies of the two types in the limit as $N\to\infty$; to 
obtain a deterministic limit, we look over timescales that are short relative to population
size. In our setting, the total population size is replaced by the local population
size, as measured by convolution with $\rho_{\epsilon}$, which we expect to be 
of order $N\epsilon^d$, and so in order to ensure a deterministic limit we 
take $\theta/(N\epsilon^d)\to 0$.
Concretely,
since $\|\rho^\epsilon_F\| \propto \epsilon^{-e/2}$,
the quadratic variation of $F(\rho^\epsilon_F\!*\!\eta(x))$
is proportional to $\theta/(N\epsilon^d)$.

@@

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa^N \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform 
	given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

>>>>

@@

By replacing $\rho$ by
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, 
we can also scale the ``width'' of the 
region over which we measure local population density. 
When the population follows~(\ref{general deterministic limit}), 
we expect that if we take a second limit of $\epsilon \to 0$,
thus scaling the kernels appearing in $r$, $\gamma$, and $F$ and making interactions pointwise,
we should recover a nonlinear PDE.
We verify that this is indeed the case in two important examples:
a special case of the porous medium equation with a logistic growth term, in which
the limiting equation takes the form 
\begin{equation}
	\label{PME1}
	\partial_t \varphi = \Delta (\varphi^2)+\varphi(1-\varphi);
\end{equation}
and a wide class of semi-linear PDEs of the form 
\begin{equation}
\label{semilinear PDE}
\partial_t \varphi = \DG^*\varphi+ \varphi F(\varphi),
\end{equation}
which includes the Fisher-KPP equation and the Allen-Cahn equation.
Equations of this form have been studied extensively
in the context of spatial ecology
(see for instance \citet{lam2023introduction} and \citet{cantrell2004spatial})
and in many other fields;
for instance, \citet{ghosh2022emergent}
derive a stochastic version of~\eqref{semilinear PDE} to describe abundances of mutant bacteria strains
along the human gut,
\citet{li/buenzli/simpson:2022} study the effects of nonlinear diffusion
on long-term survival of a lattice-based interacting particle system,
and \citet{birzu/hallatschek/korolev:2017} describe genetic variation in expanding waves
using both forwards and backwards-time arguments.
Most of this work is theoretical; for empirical studies see for instance
\citet{adler2018interspecific} or \citet{zhu2023density}.
We do not study the effect of movement of adults,
which can additionally affect the limiting equations:
see for instance \citet{holmes/lewis/banks/veit:1994} or \citet{potts/borger:2023},
especially if movement depends on population density
(as in \citet{perkins:1992} and \citet{birzu2019genetic}).
% MORE DETAILS:
% \cite{li/buenzli/simpson:2022} write down an interacting particle
% system on a hexagonal lattice (with at most one individual per site) that incorporates an effect
% of local crowding on individual motion. Their equation
% corresponds to a reaction-diffusion equation with nonlinear diffusion and a 
% bistable reaction term of the form~(\ref{Birzu nonlinearity}) and they
% investigate the interplay between nonlinear diffusion and long-term 
% survival or extinction of the population.

@@

Since $F$ is bounded above,
the final assumption that $\mu_\theta \ge 0$ --
or, equivalently, that $F(x,\eta) \le \theta r(x,\eta) \gamma(x,\eta)$ --
will always be true for large enough $\theta$ 
as long as $r$ and $\gamma$ are bounded away from zero.

@@

\begin{figure}
    \begin{center}
        % \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        % \includegraphics{figures/ex1b/fkpp_123.density.3.0}
        \includegraphics[width=2.5in]{figures/fig1a}
        \includegraphics[width=2.5in]{figures/fig1b}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha = \theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal ($q_\theta$) and interactions (here, only $\rho_F$)
        are Gaussian with standard deviation 1,
        and the equilibrium density ($N$) is 10 individuals per unit area.
        The remaining parameters are constant: $r \equiv \gamma \equiv 1$.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}

When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\citep{etheridge:2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence} when $\alpha>0$.
We do not address the question of uniqueness here,
but (as stated below in Proposition~\ref{thm:mmt_application}),
a consequence of the Markov Mapping Theorem (Theorem~\ref{thm:mmt}) is that 
uniqueness of the martingale problem for the (yet to be defined) lookdown process would imply
uniqueness of solutions to this process.

Alternatively, the Dawson--Girsanov transform could be used to show uniqueness \revpoint{1}{12}
in the special case of a superprocess with nonlinear death rates:
if $r$ and $\gamma$ only depend on $x$ (not $\eta$),
then the process with $F=0$ is a heterogeneous branching superprocess
and hence the corresponding martingale problem is unique (see Section 4.3 of \cite{dawson:1993}).
Then, the Dawson--Girsanov transform (Theorem 7.2.2 of \cite{dawson:1993},
extended to measures with nonconstant mass as in Section 10.1.2 of \cite{dawson:1993})
would provide the Radon-Nikodym derivative of the law of the process with more general $F$
relative to the law of the process with $F=0$
under suitable conditions.
% To do this may require some boundedness assumptions
% (e.g., $r\gamma$ bounded away from zero and $F$ bounded below).

In a different but related setting, the Perkins stochastic calculus
(and its adaptation to a lookdown setting \cite{donnelly/kurtz:1999})
provides uniqueness for a different but related class of processes,
in which interactions affect the dispersal mechanism
(rather than reproduction) of the superprocess \cite{perkins:1992}.
% We refer to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
% for approaches to showing that these sorts of martingale problems are well-posed.

@@

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
$\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
Furthermore, suppose that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exists $\lambda\in (0,1)$ for which
\begin{equation*}
\sup_\epsilon \int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx < \infty,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx < \infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}

@@

However, the \emph{lookdown construction} in Section~\ref{sec:lookdown}
does just this -- \llabel{whatis_lineage}
the construction enables us to recover information about ancestry
in the infinite population limit,
and thus gives a concrete meaning to $(L_t)_{t \ge 0}$. \llabel{Linf_mention}
Roughly speaking,
each particle is assigned a unique ``level'' from $[0,\infty)$
that functions as a label and thus allows reconstruction of lineages.
The key to the approach is that levels are assigned in such a way as to be exchangeable, 
so that sampling a finite number, $k$ say, of
individuals from a given 
region is equivalent to looking at the individuals in that region with the $k$ lowest levels.
Moreover, as we pass to the infinite population limit,
the collection of (individual, level) pairs converges,
as we show in Theorem~\ref{thm:lookdown_convergence}.
See~\citet{etheridge/kurtz:2019} for an introduction to these ideas.
In particular, even in the infinite
population limit, we can sample an individual
from a region (it will be the individual in that region with the lowest level) 
and trace its line of descent. 
This will allow us to calculate, for each $x$ and $y\in\IR^d$, the proportion of
the population at location $x$ in the present day population that are descended from 
a parent who was at location $y$ at time $t$ in the past. To make
sense of this in our framework, in Section~\ref{sec:lineages_proof}, 
we justify a weak reformulation of this idea.

@@

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
with mean displacement in the direction of $\grad \log(\varphi_s \gamma) - \meanq/\sigma^2$.
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

@@

Ancestry in traveling waves for populations described by reaction-diffusion equations
has been studied before by various authors, \revpoint{1}{16}
although most work assumes the diffusion term is linear
(in our notation, $r$ and $\gamma$ are constant,
but $F$ can depend on density).
For instance, following \citet{hallatschek/nelson:2008},
various authors (e.g., \citet{roques2012allee,birzu/hallatschek/korolev:2017})
describe genetic diversity
and ancestry in (possibly noisy) traveling waves for reaction-diffusion equations
in a situation that covers our first two examples below.
However, we are not aware of previous work covering the third case 
with nonlinear diffusive term (the porous medium equation).

%%%%%%%%%%
\paragraph{Fisher-KPP equation:}
Consider the classical Fisher-KPP equation,
\begin{align} \label{eqn:fkpp}
    \partial_t\varphi = \partial_{xx}\varphi + \varphi (1-\varphi) .
\end{align}
Even though we do not have an explicit formula for the wave shape in this case,
our methods provide information about ancestral lineages.
The equation has non-negative travelling wave solutions of 
speed $c$ for all $c \geq 2$, 
but, started from any compact perturbation of a Heaviside function, the 
solution will converge to the profile $w^F$ with the minimal wavespeed, $c=2$
\citep{kolmogorov/petrovsky/piscounov:1937,fife1977approach,bramson:1983}.
No matter what initial condition,
for any $t>0$ the support of the 
solution will be the whole real line. 
In this case, we must have $r = \gamma = 1$,
and $F(x, m) = 1 - m$ so $\mu_\theta(x, m) = 1 + (m-1)/\theta$.
By Corollary~\ref{cor:wavefront},
the generator of the motion of an ancestral lineage is
\begin{equation} \label{eqn:fkpp_generator}
    \Lgen_F  f
    =
    \partial_{xx}f + 2 \frac{\partial_xw^F}{w^F} \partial_x f + 2 \partial_xf .
\end{equation}
Near the tip of the wave (for $x$ large), $w^F(x) \sim e^{-x}$,
so \eqref{eqn:fkpp_generator} implies that
the motion of a lineage is close to unbiased Brownian motion.
On the other hand, in the ``bulk'', %for $x < 0$ 
a lineage behaves approximately as
Brownian motion with drift at rate two to the right.
This implies that
ancestral lineages are pushed into the tip of the wave,
and there is no stationary distribution,
so that long-term dynamics of genetic inheritance
depend on the part of the wave not well-approximated by a smooth profile,
in agreement with the previous results referred to in the Introduction.

@@

We now look at the second order term in~(\ref{generator prelimit}),
which will converge to the quadratic variation of the limiting process.
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\int
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
\eta(dx) .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $-f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
\begin{align*}
& G''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \int
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    \eta(dx)
    +
    \int
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    \eta(dx)
\right\} \\
&\qquad
= \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
	\int\left\{
		\gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta)
    \right\} \eta(dx) .
\end{align*}
Since $\int f^2(y) r(y, \eta) q_\theta(x, dy) \to f^2(x) r(x, \eta)$
and $r\gamma  + \mu_\theta = 2 r \gamma  - F / \theta \to 2 r \gamma$
as $\theta \to \infty$,
this converges to
\begin{align*}
\frac{\alpha}{2} G''(\langle f, \eta \rangle)
    \big\langle
    2 r(x, \eta) \gamma(x, \eta) f(x)^2,
        \eta(dx)
    \big\rangle .
\end{align*}

@@

\paragraph{One-step convergence:} \label{sec:one_step_heuristics} \llabel{one_step_heuristics}
In order to pass directly to a classical PDE limit
in Theorem~\ref{thm:local_convergence}
we impose the stronger condition
that $\theta/N\epsilon^d\to 0$ and also require that
$\theta\epsilon^2\to\infty$. 
Recall that in this case, we take $\rho_F^\epsilon$ 
to be a symmetric Gaussian density with variance 
$\epsilon^2$. The condition $\theta\epsilon^2\to\infty$ 
ensures that $\epsilon^2$ is large enough relative to $1/\theta$
that the regularity gained by smoothing our population density by convolution with
$\rho_\epsilon$ is preserved under the dynamics dictated by $q_{\theta}$.
To understand the first condition, note that we are aiming to obtain a 
deterministic expression for the limiting population density. 
It is helpful to think
about a classical Wright-Fisher model (with no spatial structure and just two types, say). 
We know then that if the
timescale $\theta$ is on the same order as population size $N$, we see stochastic
fluctuations in the frequencies of the two types in the limit as $N\to\infty$; to 
obtain a deterministic limit, we look over timescales that are short relative to population
size. In our setting, the total population size is replaced by the local population
size, as measured by convolution with $\rho_{\epsilon}$, which we expect to be 
of order $N\epsilon^d$, and so in order to ensure a deterministic limit we 
take $\theta/(N\epsilon^d)\to 0$.
Concretely,
since $\|\rho^\epsilon_F\| \propto \epsilon^{-d/2}$,
the quadratic variation of $\rho^\epsilon_F\!*\!\eta(x)$
is $\mathcal{O}(\theta/(N\epsilon^d))$.

@@

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa^N \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform 
	given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.
    If uniqueness holds for the $(A^N, \lp^N_0)$-martingale problem,
    then uniqueness also holds
    for the martingale problem of Definition~\ref{defn:mgale_construction}.
